The Neuro Bureau ADHD-200 Preprocessed Repository

### 2015-12-22 Bellec_ADHD200_2016.tex
additional changes

\section{Recommendations for using the data}

Besides identifying rs-fMRI based biomarkers of ADHD --- the reason that the ADHD-200 Sample was originally created and its most common use so far --- there are a variety of interesting analyses that can be performed using this data \cite{craddock2015}. With 585 TDC participants between the ages of 7 and 26, and the inclusion of measures of intelligence, the ADHD-200 Preprocessed is a valuable resource for mapping developmental trajectories \cite{dosenbach2010, zuo2010, dimartino2014} and other sources of inter-individual variation \cite{kelly2012}. Perhaps most exciting are new methods that cluster individuals based on connectivity profiles \cite{miranda2014connectotyping, finn2015fingerprinting}, which are providing new hope for using neuroimaging data to parse the heterogeneity within mental health disorders \cite{castellanos2013}. One of the biggest outstanding needs for neuroimaging, and connectomics in particular, is the development and validation of new analytical tools and processing strategies \cite{craddock2013,varoquaux2013,craddock2015} and it is our hope that the ADHD-200 Preprocessed data will become a benchmark dataset for evaluating new tools as they are proposed.

The two biggest challenges for using this data are head motion \cite{power2012, vandijk2012, sattertwhaite2012, fair2012ADHD-200, yan2013motion, yan2013small} and inter-site variation in the equipment, parameters, and experimental procedures used to collect the data \cite{Olivetti2012,yan2013std}. A variety of different approaches have been proposed for addressing head motion in hyperkinetic populations \cite{power2012, sattertwhaite2013}, and in the ADHD-200 Sample in particular \cite{fair2012ADHD-200}, that should be considered when analyzing the data. At the very least, some statistic that characterizes individual motion (such as root mean square deviation \cite{JenkinsonTR99}) should be included as a nuisance regressor in the group level model \cite{sattertwhaite2012, yan2013motion}. Differences in the manner in which data was collected at each site can introduce addition and multiplicative effects (batch effects) to the data, which may obscure the underlying biological signal \cite{Olivetti2012,yan2013std}. Including a regressor for acquisition protocol (see tables \ref{smri_params_table} and \ref{rsfmri_params_table} for different protocols), the average pairwise correlation between all regions in the brain (GCOR) \cite{saad2013gcor}, or the whole-brain average of the feature under inquiry \cite{yan2013std}, have all been shown effective for dealing with inter-site variation.

% CC-@PB please include any citations to your work that you think help in the talk of parcellations
Beyond personal ideology, or the availability of particular features, there is no clear reason to prefer one pipeline over the other. Hopefully, the presence of multiple pipelines will enable researchers to compare the robustness of their tools or analysis results to different processing choices. It isn't clear which of the parcellations provide the best time series, although there is growing consensus that functionally defined brain regions areas represent the brains intrinsic connectivity better than structurally defined atlases \cite{craddock2012whole, Blumensath2013, thirion2014}. Similarly, it isn't clear which scale of parcellation is the best. More, smaller parcels tend to be better at representing connectivity present at the voxel scale \cite{bellec2006identification, craddock2012whole, thirion2014}, but fewer parcels may be more manageable for the analysis at hand.

@@

The ADHD-200 Preprocessed initiative was a success in terms of its primary objective: the data repository was effectively used by many researchers during and after the ADHD-200 competition, with over 10,500 downloads by more than 600 users, as well as 49 publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} derived from the release in just over 3 years, with more in press or under review. The resource was also used outside of the neuroimaging community, with several publications in engineering and statistics journals that do not routinely feature neuroimaging applications \citep[e.g.][]{Liang2012, Rangarajan2015, Rangarajan2014, Mahanand2013, Ji2011, Li2015, Li2013, Liu2012, Subramanian2013, Hou2015, Deshpande2015, She2014, Chen2015, He2013, Kong2013, Yang2015, Ahn2015}. In particular, the winning team of the ADHD-200 global competition was based at the Johns Hopkins Biostatistics Department and used ADHD-200 Preprocessed to develop their diagnostic algorithm \cite{Eloyan2012}.
\par 
Our conclusions are more mitigated regarding the secondary objective of ``ADHD-200 Preprocessed'', which was to cater to multiple data processing philosophies. The release has been a success in the sense that almost all variants of processing strategies included in the ADHD-200 release have been used in follow-up published manuscripts. However, we are not aware of a publication that has performed a face-to-face comparison between processing techniques. With the approach we took, any direct comparison between Athena and NIAK was difficult to interpret, as the two pipelines did not use the same specific stereotaxic space, varied in many ways in terms of selected parameters, and also had quite different approaches for QC, derivatives or even packaging the derivatives. This lack of harmonization was essentially driven by the tight deadline for data release imposed by the competition, rather than a design choice. We since came to the conclusion that it would be useful to build the processing pipelines on a common stereotaxic space and possibly sets of registration, so that the strategies of noise reduction can be more readily compared. 
\par 
The impact of the ``ADHD-200 Preprocessed'' repository emphasized the need to reduce computational barriers to participation in discovery neuroscience, including but not limited to machine learning competitions based on neuroimaging data. This need is further emphasized by a recent survey of FCP/INDI users in which 70\% of the respondents indicated that the availability of statistical derivatives is either very important or absolutely critical to the neuroimaging community. We do not plan to expand or update the ADHD-200 release as such, which we hope will continue to serve as a legacy benchmark dataset. We still believe that a much larger-scale effort will be necessary to unlock the full potential of openly shared neuroimaging data to accelerate neuroimaging research. We are now working on new initiatives learning from the ADHD-200 Preprocessed experiment, notably including a larger number of alternative analytical workflows as well as a careful harmonization of processing, quality control and data packaging strategies. Our hope is that ADHD-200 Preprocessed and future related efforts will critically help fMRI researchers to identify optimal analytical paths for a given task.

>>>>

\section{Usage recommendations}

The ADHD-200 Sample was originally created to identify rs-fMRI based biomarkers of ADHD; however, the dataset provides ample opportunity for further analyses \cite{craddock2015}. With 585 TDC participants between the ages of 7 and 26, and the inclusion of intelligence measures, the ADHD-200 Preprocessed is a valuable resource for mapping developmental trajectories \cite{dosenbach2010, zuo2010, dimartino2014} and other sources of inter-individual variation \cite{kelly2012}. Perhaps most exciting are new methods that cluster individuals based on connectivity profiles \cite{miranda2014connectotyping, finn2015fingerprinting}, which are providing new hope for using neuroimaging data to parse the heterogeneity within mental health disorders \cite{castellanos2013}. One of the outstanding needs for neuroimaging, and connectomics in particular, is the development and validation of new analytical tools and processing strategies \cite{craddock2013,varoquaux2013,craddock2015}. In the service of this aim, the ADHD-200 Preprocessed repository has the necessary components to become a benchmark dataset for evaluating new tools as they are proposed.

The two biggest challenges for using the ADHD-200 Preprocessed data are head motion \cite{power2012, vandijk2012, sattertwhaite2012, fair2012adhd200, yan2013motion, yan2013small} and inter-site variation in the acquisition equipment, parameters, and experimental procedures \cite{Olivetti2012,yan2013std}. A variety of different approaches have been proposed for addressing head motion in hyperkinetic populations \cite{power2012, sattertwhaite2013}, and in the ADHD-200 Sample in particular \cite{fair2012adhd200}, that should be considered when analyzing the data. At the very least, some statistic that characterizes individual motion (such as root mean square deviation \cite{JenkinsonTR99}) should be included as a nuisance regressor in the group-level model \cite{sattertwhaite2012, yan2013motion}. Differences in the manner in which data was collected at each site can introduce addition and multiplicative effects (batch effects) to the data, which may obscure the underlying biological signal \cite{Olivetti2012,yan2013std}. Including a regressor for acquisition protocol (see Tables~\ref{smri_params_table} and \ref{rsfmri_params_table} for a summary of the different protocols), the average pairwise correlation between all regions in the brain (GCOR) \cite{saad2013gcor}, or the whole-brain average of the feature under inquiry \cite{yan2013std}, have all been shown to be effective for dealing with inter-site variation.

% CC-@PB please include any citations to your work that you think help in the talk of parcellations
Beyond personal ideology, or the availability of particular features, there is no clear reason to prefer one pipeline over the other. Hopefully, the presence of multiple pipelines will enable researchers to compare the robustness of their tools or analysis results to different processing choices. It isn't clear which of the parcellations provide the optimal time series, although there is growing consensus that functionally-defined brain regions represent the brain's intrinsic connectivity better than structurally-defined atlases \cite{craddock2012whole, Blumensath2013, thirion2014}. Similarly, it isn't clear which scale of parcellation is optimal. While smaller parcels tend to be better at representing connectivity present at the voxel-scale \cite{bellec2006identification, craddock2012whole, thirion2014}, fewer parcels may be more manageable for the specific analysis at hand. 
% DM: Should we also mention the degrees of freedom issue here in addition to the point regarding 'manageability'?

@@

The ADHD-200 Preprocessed initiative was successful in its primary objective: the data repository was effectively used by many researchers during and after the ADHD-200 Global Competition, with over 10,500 downloads by more than 600 users, as well as 49 resulting publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} derived from the release in just over three years, with further publications in press or under review. The resource also expanded the boundaries of the traditional neuroimaging community, with several publications in engineering and statistics journals that do not routinely feature neuroimaging applications \citep[e.g.][]{Liang2012, Rangarajan2015, Rangarajan2014, Mahanand2013, Ji2011, Li2015, Li2013, Liu2012, Subramanian2013, Hou2015, Deshpande2015, She2014, Chen2015, He2013, Kong2013, Yang2015, Ahn2015}. In particular, the winning team of the ADHD-200 Global Competition was based at the Johns Hopkins Biostatistics Department and used ADHD-200 Preprocessed to develop their diagnostic algorithm \cite{Eloyan2012}.

Our conclusions are more mitigated regarding the secondary objective of ADHD-200 Preprocessed, which was to cater to multiple data processing philosophies. The release has been a success in the sense that almost all variants of processing strategies included in the ADHD-200 Preprocessed repository have been used in follow-up publications. However, we are not aware of a publication that has performed direct comparisons between processing techniques. With the approach we took, any direct comparison between Athena and NIAK was difficult to interpret, as the two pipelines did not use the same specific stereotaxic space, varied in many ways in terms of selected parameters, and also had quite different approaches for quality control, derivatives and even packaging the derivatives. This lack of harmonization was essentially driven by the tight deadline for data release imposed by the competition, rather than a design choice. We have since concluded that it would be preferable to build the processing pipelines on a common stereotaxic space (and possibly sets of registration)
% 'sets of registration' is unclear. Perhaps: 'sets of template registration spaces'
, so that the various noise reduction strategies can be more readily compared.

The impact of the ADHD-200 Preprocessed repository demonstrated the need for reducing computational barriers to participation in discovery neuroscience, including but not limited to machine learning competitions based on neuroimaging data. This need is further emphasized by a recent survey of INDI users in which 70\% of the respondents indicated that the availability of statistical derivatives is either very important or absolutely critical to the neuroimaging community. % CC: Should we include stats or remove?
We do not plan to expand or update the ADHD-200 Preprocessed release as such, which we hope will continue to serve as a legacy benchmark dataset. We still believe that a much larger-scale effort will be necessary to unlock the full potential of openly shared neuroimaging data in the service of accelerating neuroimaging research. Current initiatives include a larger number of alternative analytical workflows as well as a careful harmonization of processing, quality control, and data packaging strategies. Our hope is that ADHD-200 Preprocessed and future related efforts will critically help fMRI researchers to identify optimal analytical paths for a given task.
